---
title: "Lab 6 Intro"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Login to the lab computer (please don't use your personal computer).  
2. Navigate to github.com and login.   
2. Use GitHub desktop to clone your repository to the desktop.   
5. Copy the class repository to the desktop (https://github.com/jmledford3115/datascibiol).  
6. Copy the files for today's lab from the class repository and paste them into **your** repository.  
7. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. What is a pipe? Why are they useful in R?
A pipe is a command in R that feeds the output from one function as the input of the next function. This helps us keep our code organized because this removes the need to have to assign new objects for every new command we want to execute.

2. What is the shortcut for making a pipe?
Control + Shift + M

3. What is the difference between `select`, `filter`, and `mutate`?
'select' is used to create a new data frame with only the variables of interest. 'filter' is used to create a new data frame that only has observations/data that fits our wanted constraints. 'mutate' is used a create a new column from existing columns of data in a data frame. 'mutate' can be used with math operations to convert units or calculate ratios, for example, in a convenient way.

4. What are the logistics of midterm 1?
It will be next Tuesday. The first 50-mins will be solo work, worth 5 points, and graded on completion/effort. The last 50-mins will be collaborative work, worth 30 points, and graded on accuracy.

### Warm-up
1. Open the data `ecs21351-sup-0003-SupplementS1.csv`
```{r}
library(tidyverse)
```
```{r}
library(janitor)
```

```{r}
ecosphere <- readr::read_csv("data/ecs21351-sup-0003-SupplementS1.csv", skip = 2)
ecosphere
```

2. Clean the names of the variables
```{r}
names(ecosphere)
```
```{r}
ecosphere <- janitor::clean_names(ecosphere)
ecosphere
```

3. Explore the data using a function of your choice
```{r}
glimpse(ecosphere)
```

4. Are there more veggie, omni, insecivore, or nectar feeders in the data?
```{r}
ecosphere %>% 
tabyl(diet) %>% 
  arrange(desc(n))
```

5. Which families are Veggie?
```{r}
ecosphere %>% 
  filter(diet=="Vegetation") %>% 
  select(family, diet) %>% 
  tabyl(diet)
```

