---
title: "CHEAT SHEET!"
output: 
  html_document: 
    keep_md: yes
date: "2023-01-31"
---
```{r}
#Load the tidyverse and janitor package!

install.packages("tidyverse")
library(tidyverse)

install.packages("janitor")
library(janitor)

install.packages("skimr")
library("skimr")
```

**FORMATTING*
#title
_italics_
**bold**

Insert website link:
[webpage name](website URL)

Insert email link:
[person name](mailto:emailaddress)`

**LAB 1**

```{r}
#Get working directory.

getwd()
```

```{r}
#Make a vector/assign objects.

x <- c(4, 6, 8, 5, 6, 7, 7, 7)
```

```{r}
#Basic Calculations

mean()
median()
sd()
```

**LAB 2**

```{r}
object_total <- sum(object_1, object_2)
```

```{r}
#Data Types

my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```

```{r}
#Specifying Kind of Data

class(my_numeric)
is.integer(my_numeric) #is my_numeric an integer?
my_integer <- as.integer(my_numeric) #create a new object specified as an integer

```

```{r}
#Missing Data

my_missing <- NA
is.na(my_missing)
anyNA(my_missing)
mean(new_vector, na.rm=T) #na.rm removes the NA values in the vector
```

```{r}
#Vectors
 
my_vector <- c(10, 20, 30) #Numeric Vector
days_of_the_week <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")     #Character Vector
my_vector_sequence <- c(1:100)

days_of_the_week[3] #pulls out 3rd data point in vector
my_vector_sequence[my_vector_sequence<10] #use operators such as <, >, ==, <==
```

#Data Matrices
```{r}
#STEP ONE: Make all the vectors.

Philosophers_Stone <- c(317.5, 657.1)
Chamber_of_Secrets <- c(261.9, 616.9)
Prisoner_of_Azkaban <- c(249.5, 547.1)
Goblet_of_Fire <- c(290.0, 606.8)
Order_of_the_Phoenix <- c(292.0, 647.8)
Half_Blood_Prince <- c(301.9, 632.4)
Deathly_Hallows_1 <- c(295.9, 664.3)
Deathly_Hallows_2 <- c(381.0, 960.5)
```

```{r}
#STEP TWO: Combine all vectors into one object.

box_office <- c(Philosophers_Stone, Chamber_of_Secrets, Prisoner_of_Azkaban, Goblet_of_Fire, Order_of_the_Phoenix, Half_Blood_Prince, Deathly_Hallows_1, Deathly_Hallows_2)
```

```{r}
#STEP THREE: Create matrix with matrix() 

harry_potter_matrix <- matrix(box_office, nrow = 8, byrow = T)
```

```{r}
#STEP FOUR: Name the rows and columns.

region <- c("US", "non-US")
titles <- c("Philosophers_Stone", "Chamber_of_Secrets", "Prisoner_of_Azkaban", "Goblet_of_Fire", "Order_of_the_Phoenix", "Half_Blood_Prince", "Deathly_Hallows_1", "Deathly_Hallows_2")

colnames(harry_potter_matrix) <- region
rownames(harry_potter_matrix) <- titles
```


```{r}
#Using Data Matrices

global <- rowSums(harry_potter_matrix)

all_harry_potter_matrix <- cbind(harry_potter_matrix, global)    #Add new column
harry_potter_matrix[2,1]                                         #Selects the value in the first column, second row.
harry_potter_matrix[1:4]                                         #Selects the specified elements in a column
non_us_earnings <- all_harry_potter_matrix[ ,2]                  #Select values in the entire 2nd column.    [ROW, COLUMN]
```

**LAB 3**

#Data Frames
```{r}
Sex <- c("male", "female", "male")    #Establish the vectors.
Length <- c(3.2, 3.7, 3.4)
Weight <- c(2.9, 4.0, 3.1)
```

```{r}
hbirds <- data.frame(Sex, Length, Weight)  #Create the data frame with data.frame().
```

```{r}
names()       #Names of variables.
dim(hbirds)   #To find dimensions!
str()
```

```{r}
#Renaming Variables

hbirds <- data.frame(sex = Sex, length_in = Length, weight_oz = Weight) 
```

```{r}
#[ROW, COLUMN]

hbirds[1,]      #1st Row
hbirds[ ,3]     #3rd Column
```

```{r}
#Select values in an entire column using the `$` sign

w <- hbirds$weight_oz    
w
mean(w)
```


```{r}
#Add a row using rbind().

new_bird <- c("female", 3.6, 3.9)  
hbirds<- rbind(hbirds, new_bird)

#Add a column,
hbirds$neighborhood <- c("lakewood", "brentwood", "lakewood", "scenic Heights")
```

```{r}
# Writing Data to File

write.csv(hbirds, "hbirds_data.csv", row.names = FALSE)
```

```{r}
#Import .csv files.

hot_springs <- readr::read_csv("hsprings_data.csv")
```

```{r}
#Change variable scientist to factor (change class)

hot_springs$scientist <- as.factor(hot_springs$scientist)
class(hot_springs$scientist)
```

```{r}
#Summary Function

summary()
glimpse()
dim()
nrow()
ncol()
names()     #names of variables
head(fish, n = 10)
tail(fish, n = 10)
max()
```


```{r}
table(fish$lakeid)
little_fish <- filter(fish, length<=100)   #Filter
```

 **LAB 4**

```{r}
#select

select()
select(fish, fish_id:length)                                        #select in between fish_id and length
select(fish, -fish_id, -annnumber, -length, -radii_length_mm)       #pulls everything except certain variables
select(fish, contains("length"))                                    #only variables with length in the name
select(fish, starts_with("radii"))                                  #what variable starts with

Options to select columns based on a specific criteria include:  
1. ends_with() = Select columns that end with a character string  
2. contains() = Select columns that contain a character string  
3. matches() = Select columns that match a regular expression  
4. one_of() = Select columns names that are from a group of names 

select_if(fish, is.numeric)
select_if(fish, ~!is.numeric(.)) #Select all columns that are *not* a class of data
```


```{r}
#filter

filter(fish, lakeid == "AL")
filter(fish, length >= 350)
filter(fish, lakeid != "AL")                       #Exclude certain observations
filter(fish, length %in% c(167, 175))              #Filtering multiple values within the same variable
filter(fish, between(scalelength, 2.5, 2.55))      #Range of specific values
filter(fish, near(radii_length_mm, 2, tol = 0.2))  #Extract observations "near" a certain value but you need to specify a tolerance.

filter(fish, lakeid == "AL" & length > 350)        #and
filter(fish, lakeid == "AL" | length > 350)        #or
```

**LAB 5**


```{r}
mammals %>% mutate_if(is.character, as.factor)      #shift + command + m     #if it is character, change it into factor
```


```{r}
#arrange

arrange(scalelength)                                #default is ascending order
arrange(desc(scalelength))                          #descending order
```


```{r}
mutate(length_mm = length*10)                       #Create a new column from existing columns in a data frame
mutate_all(tolower)                                 #Changes all entries to lowercase.
mutate(across(c("order", "family"), tolower))       #Specify individual columns
```


```{r}
#if_else()

mammals %>% 
  select(genus, species, newborn) %>%
  mutate(newborn_new = ifelse(newborn == -999.00, NA, newborn))%>%     #Replace NA placeholders with NA
  arrange(newborn)
```


```{r}
#janitor

superhero_powers <- janitor::clean_names(superhero_powers)  #clean names
tabyl(superhero_info, alignment)
```

**LAB 6**

```{r}
summary()
hist()        #histogram
skim()        #find number of unique observations within a variable
summarize()   #produce summary statistics for a given variable in a data frame
```


```{r}
#n_distinct()

msleep %>%                                  #presenting the number of distinct observations
  filter(bodywt > 100) %>% 
  summarise(n_genera=n_distinct(genus))
```


```{r}
#count()
penguins %>% 
  count(island, sort = T)                   #sort=T sorts the column in descending order
```


```{r}
#across()
penguins %>%
  summarize(across(c(species, island, sex), n_distinct))      # using `filter()` and `select() across multiple variables

penguins %>%
  group_by(sex) %>% 
  summarize(across(contains("mm"), mean, na.rm=T))

penguins %>%
  summarize(across(contains("mm"), mean, na.rm=T))
```


**LAB 7.2: FUN WITH NA'S*
```{r}
is.na(msleep)
any_na(msleep)

msleep %>% 
  summarize(number_nas = sum(is.na(msleep)))

msleep %>% 
  summarise_all(~(sum(is.na(.))))

amniota_tidy <- amniota %>%     #Replace -999 with NAs
  na_if("-999")

naniar::miss_var_summary(amniota_tidy)
```

*LAB 8.1: pivot_long()*

```{r}
#`pivot_longer`(cols, names_to, values_to)
#`cols` - Columns to pivot to longer format
#'names_to` - Name of the new column; it will contain the column names of gathered columns as values
#`values_to` - Name of the new column; it will contain the data stored in the values of gathered columns
```

*LAB 8.2: pivot_wider()*

```{r}
#`pivot_wider`(names_from, values_from)  
#`names_from` - Values in the `names_from` column will become new column names  
#`values_from` - Cell values will be taken from the `values_from` column 

heartrate2 %>% 
  separate(patient, into= c("patient", "sex"), sep = "_")
```

*LAB 9.1: ggplot*

```{r}
geom_point()                          #scatterplot
geom_bar()                            #bar plot - good for counts! (when there is only one inputted variable
geom_col()                            #bar plot - when there is 2 inputted variables

geom_jitter()                         #helps with overplotting
geom_smooth(method=lm, se=T)          #line of best fit
```

*LAB 10.1: ggplot pt2*

```{r}
geom_boxplot()

+coord_flip()
+scale_y_log10()
options(scipen = 999)                 #cancels scientific notation

labs(title = "", 
       x="",
       y="")

+theme(plot.title = element_text(size=rel(1.25), hjust=1))  
 
#adjusting the size and face of the text. `rel()`changes the relative size of the title. Adding `hjust` allows control of title position

elephants %>% 
  ggplot(aes(x=sex, fill=sex)) +      #fill colors the bars
  geom_bar()

life_history %>% 
  ggplot(aes(x=gestation, y=log10(mass), size=mass))+   #`size` adjusts the size of points relative to a continuous variable
  geom_point(na.rm = T)
```


*LAB 10.2: ggplot pt2*

```{r}
p+geom_point(aes(shape=thermoregulation, color=thermoregulation), size=1.75)
+ geom_bar(position = "dodge") 
+scale_y_continuous(labels = scales::percent)

homerange %>% 
  ggplot(aes(x = class, y = log10.mass, group = taxon)) +         #`group` to make individual box plots for each taxon
  geom_boxplot()
```

*LAB 11.1: ggplot pt.3*

```{r}
library(tidyverse)
library(RColorBrewer)
library(paletteer)
library(janitor)
library(here)
library(ggthemes)

%>% mutate(year=as_factor(year))   #change year into factor so ggplot lists ALL the years in the graph
geom_line()
geom_histogram(alpha = 0.4, color = "black", fill = "deepskyblue4", bins=40)
grDevices::colors()
geom_density(fill="deepskyblue4", alpha  =0.4, color = "black")

homerange %>% 
  ggplot(aes(x=log10.mass)) +
  geom_histogram(aes(y = ..density..), fill = "deepskyblue4", alpha = 0.4, color = "black")+
  geom_density(color = "red")+
  labs(title = "Distribution of Body Mass")                          #Combine both histogram and density.

homerange <- homerange %>% 
             mutate(mass_category = case_when(log10.mass <= 1.75 ~ "small",
                                              log10.mass > 1.75 & log10.mass <= 2.75 ~ "medium",
                                              log10.mass > 2.75 ~ "large"))
```


*LAB 11.2: ggplot pt.3*

```{r}
#Some basic themes.
theme_gray(): Gray background color and white grid lines. Put the data forward to make comparisons easy.
theme_bw(): White background and gray grid lines. May work better for presentations displayed with a projector.
theme_linedraw(): A theme with black lines of various widths on white backgrounds, reminiscent of a line drawings.
theme_light(): A theme similar to theme_linedraw() but with light grey lines and axes, to direct more attention towards the

ls("package:ggthemes")[grepl("theme_", ls("package:ggthemes"))]  #list of themes

RColorBrewer:
  
display.brewer.pal(4,"GnBu")
+`scale_colour_brewer()` is for points  
+`scale_fill_brewer()` is for fills 

p+scale_fill_brewer(palette = "BuPu")+
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 60, hjust=1))+
  labs(title = "Observations by Taxon in Homerange Data",
       x = NULL,
       y= "n",
       fill= "Trophic Guild")

colors <- paletteer::palettes_d_names
colors

my_palette <- paletteer_d("ggprism::flames")
```


```{r}
#Adjusting the x and y limits
+
  xlim(0, 4) +
  ylim(1, 6)
```


```{r}
#Faceting

+ facet_wrap(~migratory_strategy, ncol=4)    #makes a ribbon of panels
+ facet_grid(migratory_strategy~.)           #allows control over the faceted variable; it can be arranged in rows or columns.  rows~columns
+ facet_grid(diet~habitat, scales = "free_y")#allow the comparison of two categorical variables, just remember a~b where a is rows and b is columns